---
title: "read input layouts"
output: html_notebook
---

This notebook compiles the existing functions and tests for reading input layouts outside of the context of RShiny applications.

I have written drafts of RShiny application modules, but those are not relevant here. 

As of now, the main source of these functions is the dsfworld package. 

```{r}
library(tidyverse)
```

Get sample layout directories to test updated function
```{r}
SAMPLE_LAYOUTS <- fs::dir_ls("../01-sample_inputs")
```

Style the function script
```{r}
styler::style_file("../00-function_development/01-updated_functions/read_plate_layout.R")
```
Read in the script containing the function to test
```{r}
source("../00-function_development/01-updated_functions/read_plate_layout.R")
# read_plate_layout
```

Try loading a bunch of layouts
```{r}
library(tools)
library(readr)
library(readxl)
 
# SAMPLE_LAYOUTS <- fs::dir_ls("../01-sample_inputs")
# 
# for (sample in SAMPLE_LAYOUTS) {
#   print(sample)
#   try(read_plate_layout(sample))
# }

all_layouts <- lapply(SAMPLE_LAYOUTS, read_plate_layout)
```

Test lap: load the package as it currently exists!
```{r}
library(platedescribr) # load the package

SAMPLE_LAYOUTS <- fs::dir_ls("../01-sample_inputs")

lapply(SAMPLE_LAYOUTS, platedescribr::read_layout)
```



