---
title: "read input layouts"
output: html_notebook
---

This notebook compiles the existing functions and tests for reading input layouts outside of the context of RShiny applications.

I have written drafts of RShiny application modules, but those are not relevant here. 

As of now, the main source of these functions is the dsfworld package. 

Copy-pasting those functions here...

```{r}
library(tidyverse)
```
Get sample layout directories to test updated function
```{r}
SAMPLE_LAYOUTS <- fs::dir_ls("../01-sample_inputs")
```

Read in the script containing the function to test
```{r}
source("../00-function_development/01-updated_functions/read_plate_layout.R")
read_plate_layout
```

```{r}
styler::style_file("../00-function_development/01-updated_functions/read_plate_layout.R")
```

```{r}
library(tools)

tested_layouts <- lapply(SAMPLE_LAYOUTS, read_plate_layout)
tested_layouts
```

```{r}
read_plate_layout_test <- function(filepath) {
  
  # read file based on it's type
  ext <- file_ext(filepath)
  
  raw <- switch(ext,
                csv = read_csv(filepath, col_names = FALSE),
                txt = read_tsv(filepath, col_names = FALSE),
                xlsx = read_excel(filepath, col_names = FALSE),
                xls =  read_excel(filepath, col_names = FALSE)
  ) %>% base:: suppressMessages()
  
  # handle files with or without the "Type" header
  first_cell <- raw[1,1][[1]]
  out <- switch(first_cell,
                Type = raw[-1,],
                raw)
  
  # get column names
  plate_col_names <- c("variable", "row", slice(out, 1)[-c(1,2)])
  
  # convert into layout form
  out |>
    set_names(plate_col_names) |>
    filter(row %in% c(base::letters[1:16],base::LETTERS[1:16])) |>
    discard(~all(is.na(.x)))  |> # drop columns if everything is NA
    filter(if_all(everything(), ~ !is.na(.x))) |>
    mutate(across(everything(), as.character)) |> # make all character, to prevent issues in pivot
    pivot_longer(-c("variable", "row"), names_to = "column", values_to = "value") |>
    pivot_wider(names_from = "variable", values_from = "value") |>
    mutate(well = paste0(.data$row, .data$column)) |> # make well column
    unite(condition, -c("row", "column", "well"), sep = "__", remove = FALSE)  |>
    filter(if_all(everything(), ~ .x != "Empty"), # drop if all are "empty" or NA (also empty)
           if_all(everything(), ~ !is.na(.x))) |>
    mutate(across(everything(), parse_guess)) # convert likely numeric variables to numeric
}

tested_layouts <- lapply(SAMPLE_LAYOUTS, read_plate_layout_test)

tested_layouts
  
```


