---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(tidyplate)
```

Get a sample layout
```{r}
SAMPLE_LAYOUTS <- fs::dir_ls("../01-sample_inputs")
layout <- platedescribr::read_layout(SAMPLE_LAYOUTS[2])
# layout
# A tibble:48 × 7
```

```{r}
summarise_plate(layout)
summarize_plate(layout)
```


Style the function script
```{r}
styler::style_file("../00-function_development/01-updated_functions/summarise_plate.R")
```

Read in the script containing the function to test
```{r}
source("../00-function_development/01-updated_functions/summarise_plate.R")
print(summarise_plate(layouts[[9]]))
```




Create a high-level overview of plate contents, including custom notes. 

The summary includes:
- the number of unique conditions specified in the plate
- the number of wells
- the names and values of all experimental variables
- all notes on both wells and plates

See also: annotate_plate() and annotate_well()

```{r}
library(glue)

print_variable_summary <- 
  function(layout, variable_name, ...){
  
    # get stuff like "cmpd1 in 24 wells"
  var_counts <-
      layout |> 
      select(all_of(c("well", variable_name))) |> 
      distinct() |> 
      group_by(.data[[variable_name]]) |>
      tally() |> 
      mutate(msg = paste0("--- {blue ", .data[[variable_name]], 
                          "} in ", .data$n, " wells\n")) |> 
      pull(msg)
  
  # don't include variable summaries which cannot be glued
  tryCatch({var_name <- glue_col(.literal = TRUE, "{yellow {bold {variable_name}}} (type: {glue_collapse(typeof(layout[[variable_name]]))}), includes: \n {glue_col({glue_collapse(var_counts)})}\n")
  }, error = function(e) {var_name <- glue_col(.literal = TRUE, "{yellow {bold {variable_name}}} (type: {glue_collapse(typeof(layout[[variable_name]]))}) -- variable couldn't be printed. Does it contain special characters? \n")
  }, finally = function(f) {var_name <- glue_col("variable could not be printed")

           })
    
  }

print_summary <- function(layout, plate_notes = "none", ...){ ## FAILS WITH SOME SPECIAL CHARACTERS -- write something for this 
  
  # get a few important values
  n_wells <- n_distinct(layout$well)
  n_variables <- length(names(layout)[!names(layout) %in% c("well", "row", "column", "condition")])
  variables <- names(layout)[!names(layout) %in% c("well", "row", "column", "condition")]

   # get plate heading
   plate_heading <- glue_col(.literal = TRUE,"Experiment containing {green {n_wells} wells}, with {green {n_variables}} {yellow experimental variables}")

   plate_msgs <- lapply(variables, print_variable_summary, layout = layout)

   all_notes <- tryCatch(glue_col(.literal = TRUE,"{glue_collapse(plate_notes, sep = '\n ')}"), error = function(e){})

  var_summary <- glue_col(.literal = TRUE,"{plate_heading} \n{glue_collapse(plate_msgs, sep = '\n \n')} \n \n{magenta {bold Notes:}} \n{plate_notes}")

}



layouts <- lapply(SAMPLE_LAYOUTS, tidyplate::read_layout)

print(print_plate_summary(layouts[[3]]))

# still fails, probably due to some special characters...
print(print_summary(layouts[[10]] |>  select(-SMILES))) # succeeds
print(print_plate_summary(layouts[[10]])) # fails
```

```{r}
 use_crayon = FALSE


  custom_glue <- ifelse(use_crayon,
                        yes = glue::glue_col,
                        no = glue::glue)
  
  n_wells <- 1
  n_variables <-1
  
  custom_glue(.literal = TRUE, "Experiment containing {n_wells} wells, with {n_variables} experimental variables")
  
  "Experiment containing {green {n_wells} wells}, with {green {n_variables}} {yellow experimental variables}"
```





## summarize, without glue_col for saving to PDF
```{r}
summarise_plate <- function(layout, plate_notes = "none", use_crayon = FALSE, ...) {
  
  # get a few important values
  n_wells <- n_distinct(layout$well)
  n_variables <- length(names(layout)[!names(layout) %in% c("well", "row", "column", "condition")])
  variables <- names(layout)[!names(layout) %in% c("well", "row", "column", "condition")]

  # # depending on argument, change the glue function used
  # custom_glue <- ifelse(use_crayon,
  #                       yes = glue::glue_col,
  #                       no = glue::glue)
   plate_msgs <- lapply(variables, summarise_variable, layout = layout, use_crayon = use_crayon)
   
  if(use_crayon){
   # get plate heading
  plate_heading <- glue_col(.literal = TRUE, "Experiment containing {green {n_wells} wells}, with {green {n_variables}} {yellow experimental variables}")

  all_notes <- tryCatch(glue_col(.literal = TRUE, "{glue_collapse(plate_notes, sep = '\n ')}"), error = function(e) {
    "Notes could not be printed."
  })

  var_summary <- glue_col(.literal = TRUE, "{plate_heading} \n \n{magenta {bold Notes:}} \n{plate_notes} \n \n{glue_collapse(plate_msgs, sep = '\n \n')}")


  } else {
      # get plate heading
  plate_heading <- glue(.literal = TRUE, "Experiment containing {n_wells} wells, with {n_variables} experimental variables")

  all_notes <- tryCatch(glue(.literal = TRUE, "{glue_collapse(plate_notes, sep = '\n ')}"), error = function(e) {
    "Notes could not be printed."
  })

  var_summary <- glue(.literal = TRUE, "{plate_heading} \n \n NOTES: \n{plate_notes} \n \n{glue_collapse(plate_msgs, sep = '\n \n')}")
    
  }
   
     print(var_summary) 

}

summarize_plate <- summarise_plate

summarise_variable <-
  function(layout, variable_name, use_crayon = FALSE, ...) {
    # get stuff like "cmpd1 in 24 wells"
    
     var_counts <-
      layout |>
      select(all_of(c("well", variable_name))) |>
      distinct() |>
      group_by(.data[[variable_name]]) |>
      tally() 
     
    if(use_crayon){
      
     var_counts <- 
      var_counts |>
      mutate("msg" = paste0(
        "--- {blue ", .data[[variable_name]],
        "} in ", .data$n, " wells\n"
      ))
     
     } else {
       
    var_counts <- 
      var_counts |>
      mutate("msg" = paste0(
        "--- ", .data[[variable_name]],
        " in ", .data$n, " wells\n"
      ))
     
     }
     
     var_msg <- var_counts |>  pull("msg")
     
    tryCatch(
      {
        if(use_crayon){
        var_name <- glue_col(.literal = TRUE, "{yellow {bold {variable_name}}} (type: {glue_collapse(typeof(layout[[variable_name]]))}), includes: \n {glue_col({glue_collapse(var_msg)})}\n")
      } else {
         var_name <- glue(" >> {variable_name} (type: {glue_collapse(typeof(layout[[variable_name]]))}), includes: \n {glue_collapse(var_msg)}\n")
      }
        },
      error = function(e) {
         if(use_crayon){
        var_name <- glue_col(.literal = TRUE, "{yellow {bold {variable_name}}} (type: {glue_collapse(typeof(layout[[variable_name]]))}) -- variable couldn't be printed. Does it contain special characters? \n")
      } else {
       var_name <- glue(" >> {variable_name} (type: {glue_collapse(typeof(layout[[variable_name]]))}) -- variable couldn't be printed. Does it contain special characters? \n") 
      }
        },
      finally = function(f) {
        var_name <- glue_col("variable could not be printed")
      }
    )
  
  }

plate_summary <- summarise_plate(layout, use_crayon = FALSE)
```

