---
title: "R Notebook"
output: html_notebook
---

Develop the utility functions for platedescribr package

THESE AREN"T the FINAL VERSIONS OF THESE FUNCTIONS

tiny final changes may happen in the package itself. 

```{r}
library(tidyverse)
library(rlang) # for `:=`
```

extract row or column from a well name
```{r}
# non-exported helper
wells_to_rows <- function(well_names){
  gsub('[0-9]+', '', well_names)
}

# non-exported helper
wells_to_cols <- function(well_names){
  as.numeric(gsub("\\D", "", well_names))
  } 
```

add empty wells to an incomplete layout
```{r}
add_empty_wells <- function(layout,
                            plate_type = "384",
                            .well_col = "well",
                            ...) {
  
  # generate well names for full plate
  wells <- tibble("{.well_col}" := well_names(plate_type, ...)) 
  
  filled <- right_join(layout, wells, by = .well_col)
  
  # add plate row and column, if layout has these
  if ("row" %in% names(layout) & "column" %in% names(layout)) {
    filled <- filled |> 
      mutate("row" = wells_to_rows(.data[[.well_col]]), # will [[.well_col]] work in a package?
            "column" = wells_to_cols(.data[[.well_col]])) # will [[.well_col]] work in a package?
  }

  filled
}


filled_layout <- add_empty_wells(layout)
filled_layout

dplyr::tibble(a = 1, b= 2)
```

generate names of plate wells
```{r}
well_names <- function(platetype, ..., dim1 = NULL, dim2 = NULL, .leading_zeroes = FALSE, .as_factor = FALSE) {
  # function to create a string of well names for a desired plate type 
  
  # accept number or character input
  platetype <- as.character(platetype)  
  
  # guide unusable custom dimensions
  if(platetype == "custom") {
    if(is.null(dim1) | is.null(dim2)) { 
      stop("For 'custom' plate types, you must supply a vector to both `dim1` and `dim2`.")
    }
  }

  if(!platetype %in% c("384", "96", "custom")){
    stop(paste0("Unknown `.platetype` supplied: ", platetype, ". Please specify one of: '96', '384', or 'custom'."))
  }
  
  # Generate well names  --------------------------------------------------------------
    # generate all possible combinations of supplied vectors as 
    wells <- switch(platetype, # accept number or character
            "384" = expand_wells(LETTERS[1:16], c(1:24), .leading_zeroes),
            "96" = expand_wells(LETTERS[1:8], c(1:12), .leading_zeroes),
            "custom" = expand_wells(dim1, dim2), .leading_zeroes)
 
    # turn into a factor if requested. 
    if(.as_factor){ # order determined by first vector; see expand_wells() above
      wells <- factor(wells, levels = wells)
    }
  
  wells
}
```
\
```{r}
dimensions_to_wells <- function(plate_rows, plate_cols, .leading_zeroes = FALSE) {

  if(is.null(plate_rows) | is.null(plate_cols)) {
    # ensure entries are not NULL -- expand.grid will silently output a corrupt dataframe
    stop("Please supply a non-null vector to both 'plate_rows' and 'plate_cols'.")

  } else {

    expanded <- expand.grid("rows" = plate_rows, "cols" = plate_cols) |> 
      arrange(.data$rows)
      
    # add leading zeroes to columns under 10 if requested
    if(.leading_zeroes){
      expanded <- expanded |>
 mutate("cols" = if_else(.data$cols < 10,
                                true = paste0("0", as.character(.data$cols)),
                                false = as.character(.data$cols)))
    }

    out <- expanded |>
      do.call(paste0, args =  _) |>
      unique() # remove duplicate wells
  }

out
}

dimensions_to_wells(LETTERS[1:4],  c(1:4, 4 ,4), .leading_zeroes = TRUE)

expand.grid("rows" = LETTERS[1:4], "cols" = c(1:4)) |> 
      arrange(.data$rows) |> 
        mutate("cols" = if_else(.data$cols < 10,
                                true = paste0("0", as.character(.data$cols)),
                                false = as.character(.data$cols))) |>
        do.call(paste0, args =  _)
```

```{r}
# helper function -- create chr vec of all possible combinations
expand_wells <- function(dim1, dim2, .leading_zeroes = FALSE) {
    # expand.grid("cols" = dim1, "rows" = dim2)
 expand.grid("rows" = dim1, "cols" = dim2) |> 
  dplyr::arrange(.data$rows)
}

  expand_wells(LETTERS[1:4],  c(1:4), .leading_zeroes = FALSE)
```


```{r}
# helper function -- create chr vec of all possible combinations
expand_wells <- function(dim1, dim2, .leading_zeroes = FALSE) {
    # expand.grid("cols" = dim1, "rows" = dim2)
    #  do.call(paste0, dplyr::arrange(expand.grid("cols" = dim1, "rows" = dim2), "cols"))

    if(.leading_zeroes){
      expanded <- dplyr::arrange(expand.grid("cols" = dim1, "rows" = dim2), .data$rows) |>
                    mutate(rows = if_else(rows < 10,
                        true = paste0("0", as.character(.data$rows)),
                        false = as.character(.data$rows))) |>
        do.call(paste0, args =  _)

    } else {
      expanded <- do.call(paste0, dplyr::arrange(expand.grid("cols" = dim1, "rows" = dim2), .data$rows))
    }

    expanded
}

expand_wells(dim1 = LETTERS[1:4], dim2 = c(1:4), .leading_zeroes = TRUE)
 
# examples 
# expand_wells(dim1 = LETTERS[1:4], dim2 = c(1:4), .leading_zeroes = TRUE)
# # [1] "A01" "B01" "C01" "D01" "A02" "B02" "C02" "D02" "A03" "B03" "C03" "D03" "A04" "B04" "C04" "D04"
# 
# expand_wells(dim1 = LETTERS[1:4], dim2 = c(1:4), .leading_zeroes = FALSE)
# # [1] "A1" "B1" "C1" "D1" "A2" "B2" "C2" "D2" "A3" "B3" "C3" "D3" "A4" "B4" "C4" "D4"
```

tidy well names -- coerce 
```{r}
## not sure in the end that we even want this function...
tidy_wellnames <- function(wells, ...) {
  ## definition of tidy:
  
  # captialize all letters
  wells <- toupper(wells)
  
  # in normal well list
  ####$ do add this--at least throw a warning if there are unusual well names
}

# tidied_wells <- 
#   well_names("custom", dim1 = letters[1:10], dim2 = c(1:10)) |> 
# tidy_wellnames()
# 
# well_names("custom", dim1 = letters[1:10], dim2 = c(1:10))
```



##### PLATE HANDLERS
#' Wells to rows
#'
#' Given a well name (e.g. "A1"), extract the plate row ("A").
#'
#' @param well_names A vector of well names.
#'
#' @return A vector containing only the character elements of each well name.
#' @export
#'
#' @examples
#' wells_to_rows(c("A1", "A2", "B1", "B2"))
#' # [1] "A" "A" "B" "B"
#'
wells_to_rows <- function(well_names){
  gsub('[0-9]+', '', well_names)
}


#' Wells to cols
#'
#' Given a well name (e.g. "A1"), extract the plate column (1).
#'
#' @param well_names A vector containing only the character elements of each well name.
#'
#' @return A vector containing only the numeric elements of each well name.
#' @export
#'
#' @examples
#' wells_to_cols(c("A1", "A2", "B1", "B2"))
#' # [1] 1 2 1 2
#'
wells_to_cols <- function(well_names){
  as.numeric(gsub("\\D", "", well_names))
}


#' Add empty wells ### DOCUMENTATION NOT FINISHED######
#'
#' Add wells missing from the input layout. These wells are assigned `NA` for all variables.
#'
#' @param layout A layout tibble
#' @param plate_type the type of plate
#' @param .well_col A string--the name of the column that holds the well
#' @param ... Unused, for extensibility
#'
#' @importFrom dplyr tibble right_join mutate
#'
#'
#' @return the same layout, but with all wells in the chosen plate type
#' @export
add_empty_wells <- function(layout,
                            plate_type = "384",
                            .well_col = "well",
                            ...) {

  # generate well names for full plate
  wells <- tibble("{.well_col}" := well_names(plate_type, ...))

  filled <- right_join(layout, wells, by = .well_col)

  # add plate row and column, if layout has these
  if ("row" %in% names(layout) & "column" %in% names(layout)) {
    filled <- filled |>
      mutate("row" = wells_to_rows(.data[[.well_col]]), # will [[.well_col]] work in a package?
             "column" = wells_to_cols(.data[[.well_col]])) # will [[.well_col]] work in a package?
  }

  filled
}


#' Get the names of wells from a given plate  ### DOCUMENTATION NOT FINISHED######
#'
#' @param platetype type of plate
#' @param ... Unused, for extensibility
#' @param dim1 plate rows
#' @param dim2 plate cols
#' @param .leading_zeroes add leading zeroes?
#' @param .as_factor make the output a factor?
#'
#' @return a vector of well names.
#' @export
#'
#' @examples
#' well_names(platetype = 384)
#'
well_names <- function(platetype, ..., dim1 = NULL, dim2 = NULL, .leading_zeroes = FALSE, .as_factor = FALSE) {
  # function to create a string of well names for a desired plate type

  # accept number or character input
  platetype <- as.character(platetype)

  # guide unusable custom dimensions
  if(platetype == "custom") {
    if(is.null(dim1) | is.null(dim2)) {
      stop("For 'custom' plate types, you must supply a vector to both `dim1` and `dim2`.")
    }
  }

  if(!platetype %in% c("384", "96", "custom")){
    stop(paste0("Unknown `.platetype` supplied: ", platetype, ". Please specify one of: '96', '384', or 'custom'."))
  }

  # Generate well names  --------------------------------------------------------------
  # generate all possible combinations of supplied vectors as
  wells <- switch(platetype, # accept number or character
                  "384" = expand_wells(LETTERS[1:16], c(1:24), .leading_zeroes),
                  "96" = expand_wells(LETTERS[1:8], c(1:12), .leading_zeroes),
                  "custom" = expand_wells(dim1, dim2), .leading_zeroes)

  # turn into a factor if requested.
  if(.as_factor){ # order determined by first vector; see expand_wells() above
    wells <- factor(wells, levels = wells)
  }

  wells
}


